---
title: "ranemone"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ranemone}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

[ANEMONE DB]: <https://db.anemone.bio/>


## Installation

You can install the development version from [GitHub](https://github.com/):
```{r, install}
#| eval: false
# install.packages("pak")
pak::pak("heavywatal/ranemone")
```
Use the same command from time to time to update the package.


## Usage

```{r, setup}
library(ranemone)
```


## Download data

ranemone provides a function to download data files recursively from [ANEMONE DB] in the same way as instructed on the website.
Use `ranemone.directory_prefix` option to specify an existing data tree or a new destination for downloading.
```{r, prefix}
#| eval: false
options(ranemone.directory_prefix = "~/db")
ranemone::directory_prefix()
```

Authentication requires your username and a temporary password generated in [ANEMONE DB].
```{r, download}
#| eval: false
options(ranemone.user = "your-username")
options(ranemone.password = "generated-password") # not yours

ranemone::wget_recursive("dist/MiFish/ANEMONE/")
```
This command tries not to overwrite existing files.


## Read files

[ANEMONE DB] consists of many small files.
It can take time to list and read them all.
This package provides a caching mechanism to speed up the process for repeated use.
You can check and change the cache directory as follows:
```{r, cache}
#| eval: false
ranemone::cache_dir()
options(ranemone.cache_path = "~/.cache/ranemone")
ranemone::cache_dir()
```

Read all files with the given file name recursively in `directory_prefix()`.
```{r, read}
#| eval: false
sample_df = ranemone::read_tsv_xz("sample.tsv.xz")
experiment_df = ranemone::read_tsv_xz("experiment.tsv.xz")
community_df = ranemone::read_tsv_xz("community_qc3nn_target.tsv.xz")
```

The results are cached in `cache_dir()`, which makes subsequent calls faster.
```{r, cache-list}
#| eval: false
fs::dir_ls(ranemone::cache_dir())
```
